---
title: "Simulation"
author: "Simon Dovan Nguyen"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    theme: lumen
    css: assets/styles.css
---

# Set Up

## Notes

SelectorPlotFunction is wrong The non-identical errors at the end are most likely due to the breaking ties distance metric at the end. It's probably throwing away the remaining data that are less than SelectorN/not making the distance metric working \## Libr.

```{r message=FALSE, warning=FALSE}
library(MASS)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(class)
library(glmnet)
library(nnet)
library(RcppAlgos)  #TVA
library(data.table) #TVA
```

## Load Functions

```{r}
rm(list=ls())
directory = "/Users/simondn/Documents/RashomonActiveLearning/"

if(exists("directory")){
  source(paste0(directory,"Code/functions/Auxiliary/LoadFunctions.R"))
  }else if(!exists("directory")){source("Code/functions/Main/LoadFunctions.R")}

```

# Inputs

```{r}
### Input ###
# seed = 1234567890
seed = 123
N = 1000
K = 4
NClass = 2
# ClassProportion = c(3/5, 2/5)
ClassProportion = rep(1/NClass, NClass)
MeanMatrix = rep(0,K)
CovCorrVal = -0.9
TestProportion = 0.2
SelectorN = 1
InitialN = 10
ModelType = "LASSO" #Supported Types: Logistic, LASSO, Multinomial, MultinomLASSO, RandomForest
NBins = 3
```

```{r}
### Seed ###
set.seed(seed)

### Data ###
DGPResults = GenerateDataFunc(N, K, NClass, ClassProportion, CovCorrVal, NBins = NBins)
dat = DGPResults$dat
TrueBetas = DGPResults$TrueBetas
# dat$Y = as.numeric(dat$Y)
```

```{r}
CovariateList = c("X1", "X2", "X3", "X4")
RashomonParameters = list(K = K, 
                          NBins = NBins,
                          H = Inf,                           # Maximum number of pools/splits
                          R = NBins+1,                       # Bins of each arm (assume 0 exists)
                          reg = 0.1,                         # Penalty on the splits
                          theta = 2,                         # Threshold; determine relative to best model
                          inactive = 0)
```

# Rashomon Test

```{r}
# RashomonProfile =RashomonProfileFunc(dat, K, NBins, CovariateList, LabelName)
# RashomonSetNum = RashomonProfile$RashomonSetNum
# RashomonLosses = RashomonProfile$RashomonLosses
# PredictionObsModel = RashomonProfile$PredictionObsModel
# PredictionObsModelRounded = RashomonProfile$PredictionObsModelRounded
# RashomonSetTime = RashomonProfile$RashomonSetTime
# WholeSetTime = RashomonProfile$WholeSetTime
# dat = data.frame(dat)
# 
# for(col in 1:ncol(PredictionObsModel)){
#    table(PredictionObsModel[,col]) %>% print()
#    table(PredictionObsModelRounded[,col]) %>% print()
# }
```

## Algorithm

```{r}
LabelName = "YStar"
SelectorN = 1
# ModelType = "Linear"
ModelType = "Linear"
```

```{r}
SelectorType = "BreakingTies"
# iter=1
```

```{r}
SimulationFunc(dat = dat,
               LabelName,
               CovariateList,
               TestProportion = TestProportion,
               SelectorType = "Random",
               SelectorN = SelectorN,
               ModelType = ModelType,
               InitialN = InitialN,
               seed = seed) -> SimResultsRandom
```

```{r}
SimulationFunc(dat = dat,
               LabelName,
               CovariateList,
               TestProportion = TestProportion,
               SelectorType = "BreakingTies",
               SelectorN = SelectorN,
               ModelType = ModelType,
               InitialN = InitialN,
               seed = seed) -> SimResultsBreakingTies

```

## Visualize
```{r}
TestSetPredictionsRandom = SimResultsRandom$TestSetPrediction
TestSetPredictionsBreakingTies = SimResultsBreakingTies$TestSetPrediction

ModelListRandom = SimResultsRandom$ModelList
ModelListBreakingTies = SimResultsBreakingTies$ModelList
```

```{r}
### Stop Iter ###
TailN = 10
VarThreshold = 0.2
ErrorThreshold = 1.1
StopIterRandom = StoppingCriteriaFunc(SimResultsRandom, 
                                      ErrorThreshold = ErrorThreshold,
                                      VarThreshold = VarThreshold, 
                                      TailN = TailN)
StopIterBreakingTies = StoppingCriteriaFunc(SimResultsBreakingTies, 
                                            ErrorThreshold = ErrorThreshold,
                                            VarThreshold = VarThreshold, 
                                            TailN = TailN)

# StopIterRandom = StopIterRandom*SelectorN
# StopIterBreakingTies = StopIterBreakingTies*SelectorN

### Plots ###
xupper = NULL
ActiveLearningPlotFunc(Error = SimResultsRandom$Error,
                       StopIter = StopIterRandom,
                       SelectorType = SimResultsRandom$SelectorType,
                       SelectorN = SelectorN,
                       InitialTrainingSetN = SimResultsRandom$InitialTrainingSetN,
                       xupper = xupper) -> RandomPlot

ActiveLearningPlotFunc(Error = SimResultsBreakingTies$Error,
                       StopIter = StopIterBreakingTies,
                       SelectorType = SimResultsBreakingTies$SelectorType,
                       SelectorN = SelectorN,
                       InitialTrainingSetN = SimResultsBreakingTies$InitialTrainingSetN,
                       xupper = xupper) -> BreakingTiesPlot

SelectorTypeComparisonPlotFunc(SimulationType1 = SimResultsRandom,
                               SimulationType2 = SimResultsBreakingTies,
                               StopIter1 = StopIterRandom,
                               StopIter2 = StopIterBreakingTies,
                               SelectorN = SelectorN,
                               xupper = xupper) -> SelectorTypeComparisonPlot

ClassErrorPlotFunc(SimulationResults = SimResultsBreakingTies,
                   xupper = xupper) -> ClassErrorPlotBreakingTies
ClassErrorPlotFunc(SimulationResults = SimResultsRandom,
                   xupper = xupper) -> ClassErrorPlotRandom
```

```{r}
ClassErrorPlotBreakingTies
ClassErrorPlotRandom
RandomPlot
BreakingTiesPlot
SelectorTypeComparisonPlot
```

# Beta Comparison

```{r}
CoefModelListRandom = numeric(length = length(ModelListRandom) * length(TrueBetas)) %>%
  matrix(nrow = length(ModelListRandom),
         ncol = length(TrueBetas))
CoefModelListBreakingTies = CoefModelListRandom
DifferenceModelListRandom = numeric(length = nrow(CoefModelListRandom))
DifferenceModelListBreakingTies = numeric(length = nrow(CoefModelListRandom))

for(i in 1:length(ModelListRandom)){
  # CoefModelListRandom[i,] = as.numeric(coef(ModelListRandom[[i]])[-1])
  # CoefModelListBreakingTies[i,] = as.numeric(coef(ModelListBreakingTies[[i]])[-1])
  CoefModelListRandom[i,] = as.numeric(coef(ModelListRandom[[i]]))
  CoefModelListBreakingTies[i,] = as.numeric(coef(ModelListBreakingTies[[i]]))
}
```

```{r}
CovariateLook = 2
for(j in 1:length(ModelListRandom)){
  DifferenceModelListRandom[j] = abs(CoefModelListRandom[j,CovariateLook] - TrueBetas[CovariateLook])
  DifferenceModelListBreakingTies[j] = abs(CoefModelListBreakingTies[j,CovariateLook] - TrueBetas[CovariateLook])
}
```

```{r}
CovariateDiffError = data.frame(cbind(DifferenceModelListRandom,DifferenceModelListBreakingTies)) %>%
    mutate(iter = 1:length(DifferenceModelListRandom)) %>%
    pivot_longer(-iter)
colnames(CovariateDiffError) = c("iter", "SelectorType", "error")

  ### Plot ###
CovariateDiffPlot = ggplot() +
    geom_line(data = CovariateDiffError,
              mapping = aes(x = iter, y = error, color = SelectorType)) +
  ylab(paste0("X",CovariateLook, " LASSO Bias")) +
  ylim(0,1)
CovariateDiffPlot
```

```{r}
rbind(TrueBetas,
      CoefModelListRandom[nrow(CoefModelListRandom),],
      CoefModelListBreakingTies[nrow(CoefModelListBreakingTies),]) -> CoefficientResults
rownames(CoefficientResults) = c("True", "Random", "BreakingTies")
colnames(CoefficientResults) = paste0("X", 1:K)
CoefficientResults %>% round(5)
```

# Standard Error of Betas

```{r}
ModelEndBT = ModelListBreakingTies[779][[1]]
```

```{r}
SEList = numeric(4*length(ModelListBreakingTies)) %>%
  matrix(ncol = 4)

for(i in 1:length(ModelListBreakingTies)){
  SEList[i,] = summary(ModelListBreakingTies[[i]])$coefficients[,2]
}
```

```{r}
CovariateSE = SEList %>%
  data.frame %>%
  mutate(iter = SimResultsBreakingTies$InitialTrainingSetN:(nrow(SEList)+SimResultsBreakingTies$InitialTrainingSetN-1)) %>%
    pivot_longer(-iter)
colnames(CovariateSE) = c("iter", "Covariate", "SE")

  ### Plot ###
CovariateDiffPlot = ggplot() +
    geom_line(data = CovariateSE,
              mapping = aes(x = iter, y = SE, color = Covariate)) +
  ylab("Number of annotated observations")
  # xlim(700,800) +
  # ylim(0, 0.2)
CovariateDiffPlot
```
