---
title: "Rashomon Package Vignette"
author: "Garrett Allen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



# Introduction

In this vignette, we will demonstrate how to use the package rashomontva to
generate all "near-optimal" models in a factorial design setting; that is, 
models that have loss lower than a particular threshold we set. All
of the models in this package are currently simple mean models (your prediction
for an observation is the mean of the group it ends up belonging to, after 
pooling different factorial combinations) but this can be extended to work with covariates.

This package has been adapted from Apara Venkateswaran's python package `rashomon-partition-sets`,
linked here: https://github.com/AparaV/rashomon-partition-sets. 

Both packages are implementations of Venkateswaran, A., Sankar, A., Chandrasekhar, A. G., & McCormick, T. H. (2024). Robustly estimating heterogeneity in factorial data using Rashomon Partitions. arXiv preprint arXiv:2404.02141. URL: https://arxiv.org/abs/2404.02141. If you want more background on this package
and how to use it, this is a good place to start. 

This package is still in development, and this vignette will be updated as 
use changes. In particular, the section for finding the Rashomon Set for a 
single profile will likely be changed to be easier for the user in the near
future, so look out for updates to the package on Github. I will post a new vignette
whenever major features or added or how to use the package changes substansially. 

In the near term, there will be some more QOL features added to make the package
easier to use, and in the medium term, there will likely be speedups in the form
of transitioning some code to Rcpp. If you need extremely fast speed for your
project, I suggest trying to use Apara's `rashomon-partition-sets` package
n Python, as it will be significantly faster due to R being bad at performing
some key low level tasks. It is quite similar to this package, and he has
his own guide for how to use the package posted as well. 

If you have any questions or suggestions for improvement to the package please
reach out to gpa5@uw.edu, or if you're in UW Stats/Biostats, reach out to me on Slack. 

## Loading data, package, and prepping the data

First, install the package. At the moment, this must be done via devtools, but
this package will eventually be uploaded to CRAN. 

```{r installing package}
library(devtools)
devtools::install_github("Garrett-Allen/rashomon-tva-R")
```

Now that the package is installed, we want to load it:

```{r loading package}
library(rashomontva)
```

Let's also load our sample data and view it:

```{r data}
library(tidyverse)

data <- read.csv("output.csv") %>% 
  rename(ID = X) %>% 
  select(-D)
head(data)
```
From the above data, we can see that we have a factorial design structure, where
X1, X2, X3 are our covariates that each take a few discrete levels, and Y is our outcome variable.
This means that we can use this package to find the Rashomon Set!

In order to find the Rashomon Set, we will need to pass in a few parameters for our
program. Specifically, we will need to pass in the number of levels in each of our
covariates. The data can be missing factorial combinations (and the package
will still output predictions!) but for the first example, we will work
with complete data. 

Let's find out the number of levels in each of our 3 arms (X0, X1, X2).

```{r levels}
data %>% 
  count(X0)

data %>% 
  count(X1)

data %>% 
  count(X2)

```

From the above code, we see that X0 has 4 levels from 0 to 3, X1 has 3 levels
from 0 to 2, and X2 also has 3 levels. Thus, if we let R denote the vector of 
levels for each arm, R = (4,3,3) and M, the number of arms, is 3. 

# Using the Package

## Finding the Rashomon Set for a particular profile

## THIS SECTION WILL BE CHANGED TO BE EASIER SOON, BUT THIS IS HOW TO CURRENTLY DO THIS:

In the terminology of the Rashomon Set paper above, a profile refers to
which arms are active; that is, our example, the profile (0,0,1) would
refer to the factorial combinations in our data where X0 and X1 are both
not active (zero) and X2 is active (1). Due to the way the method works, 
we have to treat each profile separately when calculating the rashomon set,
then aggregate them together to get the Rashomon Set for the entire dataset. In the
following tutorial, we will demonstrate how to calculate the Rashomon Set for a
singular profile.

The key function we will use to do this is `find_rashomon_profile()`.

First, let's look at the arguments this function takes:

`data`: The dataset.
`value`: The column name of the y values- this must be a character.
`arm_cols`: A character vector containing the arm column names
`M`: Number of arms *active*
`R`: Vector of levels *for the active arms*
`H`: Maximum number of pools allowed in a model. Set to Inf if you don't care about this. 
`reg`: Regularization penalty on the number of pools: a value of 1 means
that 1 is added to the loss for each additional pool the model thinks there
is. 
`profile`: Which profile do you want to calculate the rashomon set for?
`policies`: Factorial combinations present in data for *active arms*
`policy_means`: Dataframe that gives the means for each policy in our data. 
`normalize`: Normalization factor- set this to zero if you are calculating 
this for one profile. 
`theta`: Threshold value. Models that have loss higher than this value will not
be in the Rashomon Set. Set to Inf if you want all models. 
`filtered`: Whether or not `data` is already filtered to be only observations in `profile`. 
Defaults to FALSE. 
`inactive` Level that denotes an inactive arm in data. Defaults to zero. 

While there is a lot of arguments, most are fairly easy to specify. First, we'll
set the variables that do not require speciality functions to do, and let's say
that we want to do this for profile (0,0,1). For this profile, M = 1, R = 3, 
and arm_cols = "X2"

```{r set initial parameters}
value <- "Y"
arm_cols <- c("X2")
M <- 1
R <- 3
H <- Inf
reg <- 1
profile <- c
normalize <- 0
```

Next, we calculate our policies present in the data after filtering our data
to only contain the active arms and assigning a policy and universal labels (a unique label for each
policy) to our data. 

```{r, calculate policies}

#CHECK HOW MY SUBSETTIGN WORKS!!!!!!!!!!!!!!!!!


#assigns labels and makes it easy for other functions to use: run this before
#running other key functions

#filter data to be in profile (this will be changed)
data_filtered <- data %>% 
    filter(X0 == 0, X1 == 0, X2 != 0)

#label data and prep for future steps: 
data_labeled <- rashomontva::prep_data(data_filtered, arm_cols, value, R, drop_unobserved_combinations = TRUE) %>% 
  assign_policy_label(arm_cols) %>% 
  assign_universal_label(arm_cols)

#prior two steps must be done before any of the following code

#generate policies from data
policies <- rashomontva::create_policies_from_data(data_labeled, arm_cols)

#generate policy means
policy_means <- data_labeled %>% 
  rashomontva::policy_means(value)

```

Our final step is to set theta: In practice, you should set theta by first
fitting an optimal model on your data (e.g. a lasso), finding the loss,
then setting theta to be some percentage higher than that loss, so you're 
only getting models that are sufficiently good. In our case, we will arbitrarily
set theta to be 13. 

```{r setting theta}
theta <- 13
```

The inactive level in our data is zero, so we can leave this as default, along
with filtered. We now can pass in all of our arguments. 

```{r, calculating rashomon set}

r_profile <- find_rashomon_profile(data = data_labeled,
                                   value = value,
                                   arm_cols = arm_cols,
                                   M = M,
                                   R = R,
                                   H = H,
                                   reg = reg,
                                   profile = profile,
                                   policies = policies,
                                   policy_means = policy_means,
                                   theta = theta,
                                   filtered = TRUE)
```

## Using the Rashomon Set for a profile

We now have our rashomon set for this profile. This output is a RashomonSet object 
Let's print what information we can see:

```{r}
# this print will be changed to be nicer
print(r_profile)
```
We see that our object stores 5 things, and each are lists that store this
information for each model in the Rashomon Set. Since we only have two policies
(X2 = 1, X2 = 2) we only have two possible models, both of
which happen to be in the Rashomon Set- one where both policies are in the same pool 
(so our prediction is just the mean across the entire dataset) and one where the policies are
in different pools, so that our model predicts the mean of each observation to 
be dependent on whether X2 = 1 or X2 = 2. 

We can extract the loss of each model in the RashomonSet via: 

```{r}
r_profile$losses
```
and the number of pools via:

```{r}
r_profile$num_pools
```
So, we can see that the model where we don't split (1 pools) does better than the model
where we do split (just predict the overall mean.)

The actual predictions for each model are stored in the pool_dictionary object,
which maps the universal label (as assigned by assign_universal_label) to its
predicted mean. Let's look at the pool dictionary for the second model, where
there is a split. Under the hood, this dictionary is a collections::dict()
object, and follows all the same documentation as this object. 

```{r}
dictionary <- r_profile$pool_dictionaries[[2]]

print(dictionary$as_list())
```
So we see that, for our second model, it predicts -1.92 when the universal label
is 1 (corresponding to policy (0,0,1)) and -1.97 when the universal label is 2. You
can look at data_labeled to see this correspondence between universal_label
and policy. 

Finally, if we want to predict from one of our models, we would first
process the data the same as above, then use the predict.RashomonSet function, which has arguments:

`obj`: A Rashomon Set object
`universal_label`: A vector of universal policy labels with which we want to predict. 
`model_id`: The model in the rashomon set we want to the prediction from. 

```{r predict}

#predictions from model 1
predict(r_profile, data_labeled$universal_label, model_id = 1)

#predictions from model 2
predict(r_profile, data_labeled$universal_label, model_id = 2)
```

## Calculating the Rashomon Set across the entire data (multiple profiles)

Due to the way our method works, in order to find the Rashomon Set of near
optimal models for the entire dataset, we must find the Rashomon Set separately
for each profile. Thankfully, the user does not have to deal with this issue,
as there is a function that already exists to calculate the Rashomon Set for
the entire dataset by aggregating profile Rashomon Sets. This actually means that,
if you want the Rashomon Set for each profile, you need not run the previous 
code, since this will give that output and more. 

For this section, we will use data that does not have all factorial combinations
present, so show how we can do out of sample predictions. 

The key function that does this is `aggregate_rashomon_profiles()`
Let's look at the arguments for this function:

`data`: The dataset
`value`: The column name of the y values- this must be a character.
`arm_cols`: A character vector containing the arm column names
`M`: Number of arms 
`R`: Vector of levels 
`H`: Maximum number of pools allowed in a model. Set to Inf if you don't care about this. 
`reg`: Regularization penalty on the number of pools: a value of 1 means
that 1 is added to the loss for each additional pool the model thinks there
is. 
`theta`: Threshold value. Models that have loss higher than this value will not
be in the Rashomon Set. Set to Inf if you want all models. 
`bruteforce`: Does nothing, may be updated.
`inactive` Level that denotes an inactive arm in data. Defaults to zero. 

All these arguments are relatively straightforward. Let's specify them:

```{r}
data <- read.csv("data.csv") %>% 
  select(-X)

head(data)
```
In this data, we see that there are 4 arms (X1,X2,X3,X4), each with 5 levels
(0,1,2,3,4). Some factorial combinations are missing, but this will not be an issue. 

```{r specify parameters}
value = "Y"
arm_cols = c("X1", "X2", "X3", "X4")
M = 4
R = c(5,5,5,5)
H = Inf
reg = 1
theta = 3

```

That's it! Now we can run the function:

```{r}
rset <- rashomontva::aggregate_rashomon_profiles(data = data,
                                         value = value,
                                         arm_cols = arm_cols,
                                         M = M,
                                         R = R,
                                         H = H,
                                         reg = reg,
                                         theta = theta)
```
Now we have an object rset, which is a list with two elements. The second list
`rset[[2]]` represents models that met the minimum loss to be a part of the Rashomon
Set in their respective profile. That is, each entry in `rset[[2]]` is a Rashomon Set 
object for each profile, where the threshold is the maximum loss that can occur
on that profile and still have a model in the Rashomon Set overall. 

`rset[[1]]` is a bit more complicated. The length of this list is the number
of models present in the Rashomon Set overall for this data. Each entry in
the list is itself a list, which gives the indices for model in each profile, so that
the full model is specified for the entire dataset. That is, `rset[[1]][[i]][[j]]`
gives the index of the model in the `rset[[2]][[j]]` that is a part of the ith model
in the overall Rashomon Set. 

Luckily, unless you have many, many models in the Rashomon Set, you will not need to work
with this object, and can instead use a handy function called `make_rashomon_objects(rset)`
to create a more easily interpretable list. 

```{r}
rashomon_objects <- make_rashomon_objects(rset)
```

Each object in this list is a Rashomon Set object that represents one model
in the Rashomon Set given theta. It is now easy to perform some common tasks,
like extracting losses or predicting. 

For example, to extract losses, you just need to loop over `rashomon_objects`
and sum:

```{r}
losses <- lapply(rashomon_objects, function(x) sum(as.numeric(x$losses)))

head(losses)
```
If you want predictions from the ith model in the rashomon set, this is fairly 
easy as well. There are two steps to doing this:

First, run `prep_data()` on your new data. This will add universal labels
to your data for use in the predict function, since the predict function
will only work on the policy labels. The policy labels will always
be the same for the same R, so you need not worry if you have different 
data then the data you ran `aggregate_rashomon_profiles()` with.

The arguments are self explanatory, except for `drop_unobserved_combinations`. When set to FALSE, 
the dataframe returned will contain all possible factorial combinations implied by `R`, 
including ones not present in the data, and will populate all other fields present
for these unobserved factorial combinations as `NA`. This is useful if you want
to see what the full predictions are across all factorial combinations. When set
to TRUE, it will drop these observations, and just label the data so that
you can use the `data$universal_labels` generated to make predictions. 

In our case, to demonstrate predictions on unobserved observations, I will set
this argument to FALSE. If you have testing data you just want to label, set
this argument to TRUE so that it does not populate your data with any extraneous
factorial combinations. 

```{r}
label_data <- prep_data(data, arm_cols, value, R, drop_unobserved_combinations = FALSE)
slice(label_data, 1:20)
```

Then, just run `predict.RashomonSet()` on `rashomon_objects[[i]]` to get
the predictions from the `i`th Rashomon Set. If you want to get all of the predictions,
you can just loop this function over the number of models, and store the prediction
each time you loop. 

```{r}
predictions <- predict(rashomon_objects[[5]], label_data$universal_label)
```
If there exists profiles present that have no data corresponding to them, the 
`predict` function will output NA, since there must be at least 1 data point
in a profile to have a prediction for policies in that profile. In our case, 
this means that any policy with a 0 present will have an NA output, since our
data above only has data for the profile (1,1,1,1). 

If you want to ignore these NA values, filter out policies corresponding to profiles
with no data before you run `predict()`. 

These are the main functions of the package, but you can do more: for example,
if you wanted to know the number of pools in each model of the rashomon set,
you could loop over the class variable `num_pools` as we did with the losses. Feel
free to look at the structure of a RashomonSet object to suit your own use case.

